---
title: "lab_03"
author: "Derek Willis"
date: "2023-02-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse
library(tidyverse)
# Load janitor
library(janitor)
```

## Load data and clean up the column names

Load the two datasets we worked with in the pre_lab: the Maryland payments to vendors and Maryland grants and loans and, using janitor, change the columns so that they are lower-cased and any spaces are replaced with underscores.

```{r}
# Load in the Maryland state agency payments data, making sure that it correctly sets the datatypes using all of the rows
payments <- read_csv('data/State_of_Maryland_Payments_Data__FY2008_to_FY2024.csv')
grants_loans <- read_csv('data/State_of_Maryland_Grant_and_Loan_Data__FY2009_to_FY2022.csv')
clean_names

# Load in the Maryland grants and loan data
```

## Answer questions

Q1: You want to know how much money state agencies have paid to the Prince George's County School system, but to do that you'll need to find all of the variations of how the school system might be listed. Write code to find rows where the vendor name contains the county's school system, starting with a broad search and then narrowing it down. When you have found as many as possible, save the result to a new dataframe called `pg_schools`. How many records did you find?

**A1: 66 records**

```{r}
pg_schools <- payments |> filter (
  str_detect(`Vendor Name`, "PG COUNTY PUBLIC SCHOOL") |
  str_detect(`Vendor Name`, "PRINCE GEORGES COUNTY SCHOOL") |
  str_detect(`Vendor Name`, "PRINCE GEORGES PUBLIC SCHOOLS") |
  str_detect(`Vendor Name`, "PRINCE GEORGES COUNTY PUBLIC SCHOOL") |
  str_detect(`Vendor Name`, "PRINCE GEORGES COUNTY PUBLIC SCHOOLS"))
```

Q2: Using that `pg_schools` dataframe, write code to calculate the total amount of payments to the Prince George's County School system for each fiscal year. Which fiscal year had the largest total amount? What do you notice about the years listed, and what questions do you have about that data?

**A2: 2021 had the largest total amount of payments. Every year from 2008 to 2021 isn't listed and in 2014 there was the lowest total amount of payments. I wonder why some years there were no payments and why 2014 had the least total amount of payments.**

```{r}
pg_schools |> 
group_by(`Fiscal Year`) |>
summarise(total_amount = sum(Amount)) |>
  arrange(desc(total_amount))
```

Q3. You're interested in examining the vendors paid for capital projects at the University of Maryland. Search the payments data to find those payments, then write code to produce a list of each vendor and the total amount it received. Look carefully at that list - what cleaning will it require to ensure that you have only one row per vendor and zip code? Provide at least three examples of vendor names that are similar and should be combined into one. Describe how you might clean that data so that any nearly identical vendors would look identical - what would you change and how?

**A3: We would need to make sure there are no duplicates or errors in the names of the vendors. We would also need to clean up the zip codes and make sure that there are only zip codes with 5 digits. Three examples of vendor names that are similar or the same are**

**A&I INC, A&I INC**

**AYERS SAINT GROSS, AYERS SAINT GROSS INC, AYERS SAINT GROSS INCORPORATED**

**HORD COPLAN MACHT INC, HORD COPLAN MARCHT/FREELON**

**We would use distinct () and clean_names in order to clean the data.**

```{r}
capital_projects <- payments |> 
  filter(str_detect(`Agency Name`, "CAPITAL PROJECTS- UNIVERSITY OF MARYLAND"))

```

Q4: You want to understand how much state grant money went to stem cell research projects each fiscal year. Search the data to find grants that fit that description. Once you find which columns contain the information you want, your code should standardize the data needed to find just the grants for stem cell research in a single filter. Then create a total amount of money for each fiscal year. Write a sentence summarizing how stem cell research funding has changed over time.

**A4: Since 2009, stem cell research funding has mostly decreased with some increases every now and then, especially in 2012, 2019, and 2022.**

```{r}
stem_cell_research <- grants_loans |> filter (
  str_detect(Grantor, "MD STEM CELL RESEARCH FUND") |
  str_detect(Grantor, "Maryland Technology Development Corporation/MD Stem Cell Research Fund") |
  str_detect(Grantor, "Department of Commerce - MD Stem Cell Research Fund") |
  str_detect(Grantor, "Maryland Technology Development Corporation/MD Stem Cell Research Fund, University Technology Dev. Fund") |
  str_detect(Grantor, "MARYLAND TECHNOLOGY DEVELOPMENT CORPORATION/MD STEM CELL RESEARCH FUND") |
  str_detect(Grantor, "Maryland Technology Development Corporation/Maryland Technology Transfer & Commercialization Fund")) |>

group_by(`Fiscal Year`) |>
summarise(total_amount = sum(Amount)) |>
  arrange(desc(total_amount))
```
