---
title: "lab_12"
author: "Mitchell Hang"
date: "2023-05-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

-   tidytext and our usual libraries

## Load libraries and establish settings

**Task** Create a codeblock and load appropriate packages and settings for this lab.

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse.
library(tidytext)
library(tidyverse)
library(dplyr)
library(janitor)
library(lubridate)
library(rvest)
```

## Questions

##### **Q1.** You've been assigned to report a story about the leading reasons that Maryland attorneys get sanctioned by the state for misconduct. The state [publishes lists of sanctions](https://www.courts.state.md.us/attygrievance/sanctions) that contain a short text description about the situation. Load the CSV file in the data folder containing records from fiscal year 2011 onwards. Make a list of unique words from the text column, then following the example in the pre_lab, remove common "stop words" from that list and create a list of the top 10 words containing the percentage of occurrences each word represents. What's the leading word in that answer and, broadly, what do you think the top 10 words describe?

```{r}
#Load the data
md_attorney_sanctions <- read_csv("data/md_attorney_sanctions.csv")
```

```{r}
#Create list of unique words

unique_words <- md_attorney_sanctions |> select(text) |>
  unnest_tokens(word, text)
View(unique_words)
```

```{r}
#Load stop words

data("stop_words")

stop_words <- stop_words |> 
  add_row(word = "arizona") |> 
  add_row(word = "client") |> 
  add_row(word = "funds") |>
  add_row(word = "united") |>
  add_row(word = "states") |>
  add_row(word = "colorado") |> 
  add_row(word = "records") |> 
  add_row(word = "august") |> 
  add_row(word = "year") |> 
  add_row(word = "april")  

```

```{r}
#Remove stop words from unique words list
unique_words |>
  anti_join(stop_words) |>
  group_by(word) |>
  tally(sort=TRUE) |>
  mutate(percent = (n/sum(n))*100) |>
  top_n(10)

```

**A1. The leading word is failing. I think the top 10 words have to do with the misconduct or wrongdoing that these attorneys are receiving sanctions for.**

------------------------------------------------------------------------

##### **Q2.** Let's move beyond single words to phrases. Make a list of the top 10 three-word phrases, called trigrams, based on the example from the pre_lab (you'll need to modify the example code to do this). What's the top trigram and how often does it appear? What does that phrase mean in legal terms?

```{r}
# Check the task 12 on Pre-lab 11 to see if you can use that code
trigrams <- md_attorney_sanctions |>
  unnest_tokens(trigram, text, token = "ngrams", n = 3) |>
  separate(trigram, c("word1", "word2", "word3"), sep = " ") |>
  filter(!word1 %in% stop_words$word) |>
  filter(!word2 %in% stop_words$word) |>
  mutate(trigram = paste(word1, word2, word3, sep=" ")) |>
  group_by(trigram) |>
  tally(sort=TRUE) |>
  mutate(percent = (n/sum(n))*100) |>
  top_n(10)
```

**A2. The top trigram is "attorney trust account". This is a bank account the lawyers use to hold client funds separate from the law firm's operating funds.**

------------------------------------------------------------------------

##### **Q3.** Let's drop back down to more traditional text analysis - take the top trigram from Q2 and write code to see how many times it occurs in the text column in each fiscal year. What do you think the answer produced by your code suggests? What else could you do to try and clarify the most important reasons attorneys get sanctioned?

```{r}
trigram <- "attorney trust account"
sanctions_with_phrases <- md_attorney_sanctions |> 
  mutate(trigram_occurrences = str_count(text, fixed(trigram))) |>
  group_by(fiscal_year) |>
  summarise(total_occurrences = sum(trigram_occurrences, na.rm = TRUE))
```

**A3. The results suggests that 2022 had the highest amount of lawyers being sanctioned for misuse of client funds or attorney trust accounts. Another way to try and clarify the main reasons attorneys are sanctioned would be a graph or visualization that shows the top 10 trigrams for each fiscal year.**
